---
output: pdf_document
---

# Andrew Moore, `r format(Sys.Date(), "%m/%d/%Y")`

## MATH-471, Homework 3.1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r init}
set.seed(20210922)

library(tidyverse)
library(ggtext)

theme_set(
  theme_minimal(base_size = 15) +
    theme(panel.grid.minor = element_blank())
)
```

### 1. Generate samples of n = 100 for 10 different distributions.

```{r draw samples, echo = TRUE}
# samples from each distribution stored in a list
samples <- list(
  "***A.*** Uniform"     = runif(100, 2, 5), 
  "***B.*** Binomial"    = rbinom(100, 25, 0.1), 
  "***C.*** Exponential" = rexp(100, 10), 
  "***D.*** Gamma"       = rgamma(100, 8, 10), 
  "***E.*** T"           = rt(100, 25), 
  "***F.*** F"           = rf(100, 15, 15), 
  "***G.*** Chi-square"  = rchisq(100, 9), 
  "***H.*** Weibull"     = rweibull(100, 1, 1.5), 
  "***I.*** Cauchy"      = rcauchy(100, 0, 1), 
  "***J.*** Normal"      = rnorm(100, 0, 1)
)
```

\newpage

### 2. Create histograms or density plots for each of the 10 distributions, and describe their shapes.

```{r q2, fig.width=10, fig.height=7}
enframe(samples) %>%
  unnest(value) %>%
  ggplot(aes(x = value)) +
  geom_histogram(aes(y = stat(density)), color = "white", fill = "grey") +
  geom_density(color = "orange") +
  # scale_x_continuous(breaks = 1, labels = "0") +
  facet_wrap(~name, scales = "free") +
  labs(x = "", y = "") +
  theme(strip.text = element_markdown())
```

- **A.** The shape is generally flat across the range. It is difficult to discern a natural center among the data values.
- **B.** The data look discrete, and the shape is triangular, with a peak at 2.
- **C.** The distribution is positively skewed, with a sharp drop-off of data values starting close to 0.1.
- **D.** The shape appears roughly symmetric, centered at roughly 0.75.
- **E.** The shape appears roughly symmetric, centered at 0, with a slight negative skew.
- **F.** The shape is mound-like, centered at 1, with a positive skew.
- **G.** There is moderate positive skew exhibited in the data; there appear to be very few outliers. It is similar in shape to **F.**
- **H.** The data are positively skewed, with a sharp drop-off of values starting at 2. It is similar in shape to plot **C.**
- **I.** The data seem to be centered at 0, but there is an extreme negative outlier within the sample.
- **J.** The shape is mound-like, symmetric, and centered at 0. It is similar in shape to plot **E.**


\newpage

### 3. Use QQ-plots to determine if the sample comes from a *normal* distribution.

```{r q3, fig.width=10, fig.height=7}
enframe(samples) %>%
  unnest(value) %>%
  ggplot(aes(sample = value)) +
  geom_qq_line(lty = "dashed", color = "black") +
  geom_qq(color = "orange") +
  facet_wrap(~name, scales = "free") +
  theme(strip.text = element_markdown()) +
  labs(x = "", y = "")
```

- **A.**
- **B.**
- **C.**
- **D.**
- **E.**
- **F.**
- **G.**
- **H.**
- **I.**
- **J.**
